---
title: "SQL Practice with WAI Database"
author: Keyron Linarez
execute: 
  warning: false
  message: true
---

For this project, I will be manipulating the Wideband acoustic immittance SQL data base to make some observations on their results. I will practice by re-creating the graph displayed on their website, which 12 of their data sets graphed according to frequency. I will then pick a data set and compare the results between sex. Here is their official website/citation: doi.org/10.35482/egr.001.2022

```{r}
#| echo: true
library(tidyverse)
library(RMariaDB)
con_wai <- dbConnect(
  MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")

```

```{sql}
#| connection: con_wai
#| output.var: "new_table"

SELECT Identifier, Frequency, LOG10(Frequency) AS logFreq, AVG(Absorbance) AS avgAbsorb
FROM Measurements
WHERE Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015", "Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013", "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010") AND Frequency < 8000 AND Frequency > 200
GROUP BY Frequency, Identifier
```

```{sql}
#| connection: con_wai
#| output.var: "final_table"

# combining two different data sets
SELECT p.Identifier, p.AuthorsShortList, p.Year, COUNT(DISTINCT SubjectNumber, Ear) as N_ears, AVG(Absorbance) AS avgAbsorb, Frequency, 
 CONCAT(p.AuthorsShortList, " (", p.Year, ") N=", COUNT(DISTINCT SubjectNumber, Ear), "; ") AS label
  FROM PI_Info as p
LEFT JOIN Measurements as m ON m.Identifier = p.Identifier
WHERE m.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015", "Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013", "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010") AND Frequency < 8000 AND Frequency > 200
GROUP BY Frequency, Identifier

```

```{r}
ggplot(final_table, aes(x = Frequency, y = avgAbsorb, color = label)) +
  geom_line(size = 1) +  # Use geom_line for lines
  labs(
    title = "Mean Absorbance from each publication in WAI database",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = ""
  ) +
scale_x_log10(breaks = c(200, 400, 600, 800, 1000, 2000, 4000, 6000, 8000))+ 
scale_y_continuous(breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  theme(
    axis.text.x = element_text(size = 8),  
    axis.text.y = element_text(size = 8) 
  )
```

This graph is a re-creation of the author's frequency graph, which has 12 specefic papers results mapped out on a log scale.

```{sql}
#| connection: con_wai
#| output.var: "subject_table"
# if weird data -> facet_wrap(~sex/race/etc)
SELECT Identifier, Frequency, LOG10(Frequency) AS logFreq, AVG(Absorbance) AS avgAbsorb
FROM Measurements
WHERE Identifier IN ("Shaver_2013")
GROUP BY Frequency, Identifier

```

```{sql}
#| connection: con_wai
#| output.var: "final_subjects"


SELECT s.SubjectNumber, s.Sex as sex, AVG(Absorbance) AS avgAbsorb, m.Frequency
FROM Subjects as s
LEFT JOIN Measurements as m ON m.Identifier = s.Identifier
WHERE m.Identifier IN ("Shaver_2013")
GROUP BY s.SubjectNumber, s.Sex, m.Frequency;

```

```{r}
ggplot(final_subjects, aes(x = Frequency, y = avgAbsorb, color = sex)) +
  geom_line(size = 1) +  # Use geom_line for lines
  labs(
    title = "Mean Absorbance from Shaver 2013",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
  ) +
scale_x_log10(breaks = c(200, 400, 600, 800, 1000, 2000, 4000, 6000, 8000))+ 
scale_y_continuous(breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  theme(
    axis.text.x = element_text(size = 6),  
    axis.text.y = element_text(size = 8) 
  ) + facet_wrap(~ sex)

# terminate our sql connection!
DBI::dbDisconnect(con_wai, shutdown = TRUE)

```

These two graphs show the differences betwen mean absorption between males and females measured from the Shaver 2013 study. As you can see, not much difference was recorded.

Thank you!
